
\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix,epsfig,endnotes}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[skip=0pt]{subcaption}
\usepackage[colorlinks=true,allcolors=blue,breaklinks]{hyperref}   % hyperlinks, including DOIs and URLs in bibliography

\newcommand{\Idit}[1]{{\bf [[#1]]}}
\newcommand{\inote}[1]{}
\newcommand{\remove}[1]{}
\newcommand{\comment}[1]{} 
\newcommand{\sys}{Omid} 
\newcommand{\figw}{0.925\columnwidth}

\begin{document}

\date{}

\title{\Large \bf \sys, Reloaded: Scalable and Highly-Available Transaction Processing}

\author{
{\rm Ohad Shacham}\\
	Yahoo Research
	\and	
\and
{\rm Francisco Perez-Sorrosal}\\
	Yahoo
\and
{\rm Edward Bortnikov}\\
	Yahoo Research
	\and
{\rm Eshcar Hillel}\\
	Yahoo Research
	\and
{\rm Idit Keidar}\\
	Technion and Yahoo Research
	\and
{\rm Ivan Kelly}\\
	Midokura
\and
{\rm Matthieu Morel}\\
	Skyscanner
\and
{\rm Sameer Paranjpye}\\
	Arimo
} % end author


\maketitle

\subsection*{Abstract}

We present \sys\ -- a transaction processing service that powers web-scale 
production systems at Yahoo. \sys\ provides ACID transaction semantics on top of traditional 
key-value storage; its implementation over Apache HBase is open sourced as part of Apache Incubator. 
\sys\ can
%is highly scalable. It 
serve hundreds of thousands of transactions per second
on standard mid-range hardware, while incurring minimal impact 
on the speed of  data access  in the underlying key-value store. 
Additionally, as expected from always-on production services, \sys\ is highly available. 

\sys's database-neutral philosophy and real-world deployment considerations have
led to unique design choices, including 
managing (and in particular, scaling) compute and metadata storage of transaction management independently,
and a novel high availability algorithm that runs  synchronization-free most of the time.

%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
\label{sec:intro}
\input{intro}

\section{Design Principles and Architecture}
\label{sec:design}
\input{design}

\input{relatedwork}

\section{Service Semantics and Interface}
\label{sec:semantics}
\input{semantics}

\section{Transaction Processing}
\label{sec:protocol}
\input{protocol}

\section{High Availability}
\label{sec:ha}
\input{ha}

\section{Evaluation}
\label{sec:eval}
\input{eval}

\section{Lessons Learned and Future Work}
\label{sec:lessons}

\input{lessons}


\subsection*{Acknowledgments}

We acknowledge the many contributions to Omid, as concept and code, since its early days. 
Our thanks go to Aran Bergman, Daniel Gomez Ferro, Yonatan Gottesman, Flavio Junqueira, Igor Katkov, Francis Christopher Liu, 
Ralph Rabbat, Benjamin (Ben) Reed, Kostas Tsioutsiouliklis, Maysam Yabandeh, and the Sieve team at Yahoo.  We also thank 
Dahlia Malkhi and the FAST reviewers for insightful comments.


%%%%%%%%%%%%%%%%%%%%%
%\clearpage
\small
{\bibliographystyle{acm}
\bibliography{refs}}

%%\theendnotes


\end{document}
